---
title: "Getting Started with monitOS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with monitOS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is monitOS?

**monitOS** provides statistical guidelines for monitoring overall survival (OS) 
in pivotal clinical trials for **indolent cancers** — slow-progressing diseases 
(e.g. follicular lymphoma, CLL, low-grade glioma) where experimental therapies 
are often evaluated against a watch-and-wait or standard-of-care control.

The core problem: in indolent cancers, OS benefit is hard to prove in a 
reasonable timeframe, yet OS harm must not be obscured by surrogate endpoints. 
The package implements the framework of 
[Fleming et al. (2024)](https://www.tandfonline.com/doi/full/10.1080/19466315.2024.2365648),
which defines a sequence of **positivity thresholds** — OS hazard ratio (HR) 
boundaries that must be met at each planned analysis to give sufficient 
reassurance that the experimental treatment does not cause unacceptable harm.

> **Key convention**: HR < 1 indicates benefit; HR ≥ 1 indicates potential harm.

---

## Core parameters

Three clinically-determined ingredients drive every monitoring plan:

| Parameter | Argument | Meaning |
|---|---|---|
| δ_null | `hr_null` | The OS HR that is **unacceptably harmful** (e.g. 1.3 = 30% excess mortality) |
| δ_alt  | `hr_alt`  | The OS HR that is **plausibly beneficial** given the mechanism of action (e.g. 0.80) |
| Events | `events`  | Cumulative OS deaths at each planned analysis |

Two error-rate parameters complete the design:

| Parameter | Argument | Meaning |
|---|---|---|
| γ_FA | `falsepos` | One-sided false positive rate tolerated at the **Final Analysis** (typically 0.025) |
| β_PA | `power_int` | Statistical power at the **Primary (interim) Analysis** under δ_alt (typically 0.9) |

---

## Installation

```{r, eval=FALSE}
install.packages("monitOS")
```

```{r setup}
library(monitOS)
```

---

## Scenario 1 — Standard trial: Follicular Lymphoma

### Clinical context

A Phase 3 randomised trial evaluates a novel PI3Kδ inhibitor combined with 
standard immunochemotherapy (R-CHOP) versus R-CHOP alone in previously 
untreated Grade 1–2 follicular lymphoma — a classic indolent B-cell malignancy 
with a long natural history.

- **Randomisation**: 1:1 (experimental : control)  
- **δ_null = 1.30** — a 30% increase in mortality risk is deemed unacceptable 
  for a population where watch-and-wait is a safe alternative  
- **δ_alt = 0.80** — a 20% reduction in mortality risk is plausible based on 
  the mechanism and Phase 2 PFS data  
- **5 analyses** planned at 50, 80, 110, 140, and 175 OS events  
- **γ_FA = 0.025**, **β_PA = 0.90**

```{r scenario-1}
fl_trial <- bounds(
  events     = c(50, 80, 110, 140, 175),
  power_int  = 0.90,   # 90% power at primary analysis under δ_alt
  falsepos   = 0.025,  # one-sided α at final analysis
  hr_null    = 1.30,   # unacceptable OS harm threshold
  hr_alt     = 0.80,   # plausible OS benefit
  rand_ratio = 1       # 1:1 randomisation
)

fl_trial$summary
```

### Interpreting the output

Each row corresponds to one planned analysis. The key columns are:

- **OS HR threshold for positivity**: if the observed OS HR at that analysis is 
  *at or below* this value, the treatment is considered reassuring (it does not 
  appear to cause unacceptable harm). Notice thresholds tighten as evidence 
  accumulates — from a lenient ~1.13 at 50 events down to ~0.97 at the Final 
  Analysis.

- **One-sided false positive error rate**: the cumulative Type I error spent at 
  each analysis. Only the Final Analysis exhausts the full γ_FA = 0.025.

- **Posterior probability HR ≥ δ_null**: the Bayesian probability that the 
  *true* OS HR exceeds the unacceptable threshold, given the observed estimate 
  equals the positivity boundary — a useful decision-support metric alongside 
  the threshold itself.

- **Predictive probability (%)**: at each interim, the probability that the 
  Final Analysis will also meet its positivity threshold, *assuming the current 
  trend continues*. This helps early stopping discussions.

---

## Scenario 2 — Edge case: CLL Maintenance with Unequal Randomisation

### Clinical context

A Phase 3 trial evaluates a BCL-2 inhibitor maintenance strategy versus 
observation in patients with chronic lymphocytic leukaemia (CLL) who achieved 
response to first-line chemo-immunotherapy.

This scenario introduces two complications common in real drug development:

1. **2:1 randomisation** — more patients are allocated to the experimental arm 
   to accelerate safety and biomarker data collection.
2. **Marginal benefit sensitivity analysis** — the sponsor wishes to understand 
   monitoring performance under a *pessimistic* scenario where the drug provides 
   only a modest 5% reduction in mortality (HR = 0.95), well below the expected 
   HR = 0.80 used to size the trial.

- **δ_null = 1.25** — a tighter margin than Scenario 1 because CLL patients 
  already have a relatively good prognosis; even modest harm is unacceptable  
- **δ_alt = 0.80** — 20% OS improvement anticipated based on PFS phase 2 data  
- **5 analyses** at 60, 95, 130, 165, and 200 OS events  
- **hr_marg_benefit = 0.95** — the marginal benefit scenario to stress-test

```{r scenario-2}
cll_trial <- bounds(
  events          = c(60, 95, 130, 165, 200),
  power_int       = 0.90,
  falsepos        = 0.025,
  hr_null         = 1.25,   # tighter harm margin for CLL
  hr_alt          = 0.80,
  rand_ratio      = 2,      # 2:1 experimental : control
  hr_marg_benefit = 0.95    # sensitivity: only 5% OS benefit
)

cll_trial$summary
```

### Key observations for the edge case

**Effect of unequal randomisation**: The 2:1 allocation reduces the Fisher 
information per event compared to 1:1, requiring slightly higher observed 
thresholds (more room for the HR to be above 1 while still passing). This 
is reflected in the positivity thresholds being fractionally higher than in 
Scenario 1 at equivalent event counts.

**Marginal benefit column**: The rightmost column shows the probability of 
meeting each positivity threshold if the true OS HR is only 0.95 — a scenario 
where the drug barely works. Even under this pessimistic assumption the 
monitoring thresholds are likely to be met at most analyses, providing 
reassurance that the monitoring plan does not require large OS benefit to 
conclude "no unacceptable harm."

---

## Quick reference

```{r quick-ref, eval=FALSE}
# Minimal call — all other arguments have sensible defaults
bounds(events = c(60, 100, 150))

# Full call
bounds(
  events          = c(60, 89, 110, 131, 178),
  power_int       = 0.90,
  falsepos        = 0.025,
  hr_null         = 1.30,
  hr_alt          = 0.80,
  rand_ratio      = 1,
  hr_marg_benefit = NULL   # omit or set to a scalar HR for sensitivity analysis
)
```

For an interactive version of these analyses, visit the 
[Shiny app](https://opensource.nibr.com/monitOS/shinylive-app/) — no R 
installation required.

